---
title: "Homework 3"
author: Diana Sanchez
output: github_document
---

This is my solution to HW3.

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(p8105.datasets)
knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

Observations are the level of items in orders by user. There are user / order variables (user ID, order ID, order day, and order hour). There are also item variables (name, aisle, department, and some numeric codes). 


How many aisles, and which are most items from? The following code counts the number of times an aisle is listed and orders it from largest to smallest number.

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```

There are 134 aisles and the one where most items are from is fresh vegetables.


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

The following code creates a plot, organizes the x-axis labels, reorders the x-axis points according to n from smallest to largest.

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>%
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	) %>%
knitr::kable()
```

## Problem 2

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel_df =
  read_csv(
    "./data/accel_data.csv") %>%
  pivot_longer(
    activity.1:activity.1440,
    names_to = "minute",
    names_prefix = "activity.",
    values_to = "activity_count"
  ) %>%
  group_by(day) %>%
  mutate(
    day_type = case_when(
    day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday",
    day %in% c("Saturday","Sunday") ~ "weekend")) %>%
  mutate(
    day = factor(day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  mutate(
    week = factor(week, levels = c(1:5)))
```

This dataset contains `r nrow(accel_df)` rows/observations and `r ncol(accel_df)` columns. Variables in this data set include week, day_id, day, minute, activity_count, and day_type.


Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_df %>%
  group_by(day_id, week, day) %>%
  summarise(
    total_activity = (sum(activity_count))) %>%
  arrange(day) %>%
  arrange(week) %>%
  knitr::kable()
```

I cannot see a clear overall trend, but I did notice that the two lowest total activity counts were both on Saturdays.


Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_df %>%
ggplot(aes(x = minute, y = activity_count, color = day_id), aplpha= 0.5) + 
	geom_line() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


## Problem 3

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?


Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?


Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
